---
title: \Large Excercise 5 - Model Predictive Control using a Grey Box Model
author: "Marco Hernandez Velasco"
date: "August 2018"
output: 
  pdf_document:
    citation_package: natbib
header-includes:
  - \usepackage{color}
bibliography: [references.bib, packages.bib]
---
```{r setup_bibliography, include=FALSE}
# automatically create a bib database for R packages
rm(list=ls()) # remove all variables in memory

knitr::write_bib(c(.packages(),  #make a bibliography for the packages
              #packages used that will be cited
                #basic packages used for Rmarkdown
                'knitr', 'rmarkdown',  
                #actual packages used  to solve
                'tidyverse', 'ctsmr', 'expm', 'scales'),   
                'packages.bib') #file num to which the references will be saved to

#define other general chunk settings
knitr::opts_chunk$set(echo = TRUE, # show source code chunks in the output file
                      fig.height = 4, #figures height for all chunks
                      fig.align = "center") 
```
@* `r #includes all references in file, even those not mentioned in the text `

```{r initialize, message = FALSE, echo = FALSE}
library(readxl) # to import Excel files
library(tidyverse) # to use tidy data
library(ctsmr)
library(expm)
library(scales)
library(lpSolve)

#point to the functions files
source("../functions/asHours.R")
source("../functions/asP.R")
source("../functions/prbs.R")

```
##Introduction  

In this exercise we will illustrate the main concepts of **Model Predictive Control (MPC)** by using it to control the temperature of a building. As the name suggests MPC relies on models and predictions, and so before designing a controller one needs to model the system and acquire relevant forecasts. In this exercise we will focus on the control part, assuming that we have perfect forecasts and models.



```{r model, echo=FALSE, warning=FALSE, results='hide'}
# Fit grey box model ------------------------------------------------------
data <- read.csv("../data/data.csv",stringsAsFactors=FALSE)
## Give names to the series
names(data) <- c("timedate","Y1","Y2","Ta","Gv","Ph1","Ph2")

data$timedate <- asP(data$timedate)
## timedate is the time in POSIXct, make t in hours since begining
data$t <- asHours(data$timedate-data$timedate[1])

##Model
model1 <- ctsm()

model1$addSystem(dT1a ~ ( 1/(C1a*R1m)*(T1m-T1a) + 1/(C1a*R1a)*(Ta-T1a) + 1/C1a*Ph1 + p1*A1w/C1a*Gv)*dt + exp(p1a)/C1a*dw1a)

model1$addSystem(dT1m ~ (1/(C1m*R1m)*(T1a-T1m)+(1-p1)*A1w/C1m*Gv)*dt + exp(p1m)/C1m*dw1m)

## Set the names of the inputs
model1$addInput(Ta,Ph1,Gv)
## Set the observation equation: Ti is the state, Yi is the measured output
model1$addObs(Y1 ~ T1a)

## Set the variance of the measurement error
model1$setVariance(Y1 ~ exp(e11))

## Set the initial value (for the optimization) of the value of the state at the starting time point
model1$setParameter(  T1a0 = c(init=25  ,lb=0     ,ub=35 ) )
model1$setParameter(  T1m0 = c(init=25  ,lb=0     ,ub=35 ) )

## Set the initial value for the optimization
model1$setParameter( C1a = c(init=6   ,lb=1E-5  ,ub=20) )
model1$setParameter( C1m = c(init=12   ,lb=1  ,ub=50) )
model1$setParameter( R1a = c(init=10   ,lb=1     ,ub=80) )
model1$setParameter( R1m = c(init=1   ,lb=1E-10     ,ub=10) )
model1$setParameter( A1w = c(init=1   ,lb=1E-10     ,ub=10) )
model1$setParameter( p1 = c(init=0.5   ,lb=0     ,ub=1) )
model1$setParameter( p1a = c(init=1   ,lb=-30   ,ub=10 ) )
model1$setParameter( p1m = c(init=1   ,lb=-30   ,ub=10 ) )
model1$setParameter( e11 = c(init=-1  ,lb=-50   ,ub=10 ) )

## Run the parameter optimization
fit1 <- model1$estimate(data)

Estimated <- list2env(as.list(fit1$xm))

A <- matrix(sapply(fit1$model$sys.eqs$amat,
                   eval,envir = Estimated),
            nrow=length(fit1$model$states))

B <- matrix(sapply(fit1$model$sys.eqs$bmat,
                   eval,envir = Estimated),
            nrow=length(fit1$model$states))

C <- matrix(sapply(fit1$model$obs.eqs$cmat,
                   eval,envir = Estimated),
            nrow=length(fit1$model$outputs))

D <- matrix(sapply(fit1$model$obs.eqs$dmat,
                   eval,envir = Estimated),
            nrow=length(fit1$model$outputs))


```

# Question 1

How would you change the model, if instead the building was heated by floor heating?

*The effect of the radiator $\phi$ has to be set to 0 for dTi equation and added to the dTm. In the code this is done by defining an alternative B mattrix ("altB") where the first element is 0 and the other is the coefficient.*

```{r q1, echo=FALSE, warning=FALSE, results='hide'}
# Change from air heating to floor heating
altB <- B
altB[1,2] <- 0
altB[2,2] <- 1/fit1$xm[5]
```

# Question 2
Assume that you were to control the temperature of the air in this building. Would you expect it to be equally difficult for floor heating and air heating?  
*No, there is a different in the thermal mass and the time constant of the floor is higher so it takes longer to reach the desired temperature.*

If not, which one would be the most difficult and why?
*It is more difficult to control floor heating, specially if you are want sudden/fast changes. The floor heating will hold the temperature longer due to the thermal mass.*

# Question 3
Have a look at the values of A and B, and see they fit with your intuition.  

```{r q3_AB, echo = TRUE, warning = FALSE}

# Parameters of continous time model --------------------------------------
A   # System dynamics, parameters of the states
B    # Air heating, parameters of the inputs
altB # floor heating

```
```{r q3_other, echo=FALSE, warning=FALSE, results='hide'}
C
D
# Discretize Model --------------------------------------------------------
dt <- diff(data$t)[1] #all.equal(diff(diff(data$t)),rep(0,length(data$t)-2)) ## Make sure that time steps are uniform

Ad <- expm(A*dt)
Bd <- solve(A)%*%(Ad-diag(dim(A)[1]))%*%B
altBd <- solve(A)%*%(Ad-diag(dim(A)[1]))%*%altB
Cd <- C
Dd <- D

# Parameters of Discrete time model ---------------------------------------
Ad
Bd    # Air heating
altBd # Floot heating

Cd
Dd

```

# Question 4
In this question we will formulate the equations of the MPC problem. The whole point of the control is to keep the temperature within comfort boundaries, so let us start with this. Assume that we do not care about the exact temperature of the building, as long as it is between 23C and 25C.  
Now look at the data concerning the heating of the building *(data$Ph1)*. What is the maximum amount and minimum amount of heating? 

```{r q4, echo=FALSE, warning=FALSE, results='hide'}

# Heating data 
min(data$Ph1)
max(data$Ph1)

```
Finish the MPC formulation and explain it in words:  

*The limits for the set temperature are between $23C < CXt < 25C$, the limits for the input heating are between $0kW < CXt < 1.5kW$.*  
*In this model we are assuming that the heater can be turned instantly on/off but in reality you would normally add a cost to switching it in order to do the optimization.*

# Question 5

Run the section *Perform Control* of the R-script, which simulates the controlled temperature using the same outside temperature and solar radiation as the data used for fitting the grey box model. The heating is optimized by solving the problem formulated in the previous question, with comfort boundaries and constraints on the heating equipment specified initially. By default the noise of the model is turned off, so the controller is able to perfectly align the air temperature with the lower
comfort comfort boundary for maximum efficiency. 

The control horizon is specified by N and is equal to 30 by default, meaning that the controller looks 30 time steps (300 min or 5 hours) ahead in time. What happens as you vary N?

Change the variable Air to FALSE, to simulate the building with floor heating instead of air heating. How does this affect the control?

```{r q5, echo=FALSE, warning=FALSE, results='hide'}

# Perform Control --------------------------------------------------------
set.seed(133)

Controllable <- c(FALSE,TRUE,FALSE) # The controllable inputs e.g. heating in this example, but not outside air temperature or solar radiation
MaxHeat <- 1.5 # Maximum amount of heating
N <- 60 # Control horizon
Ymax <- 25 # Maximum temperature
Ymin <- 22.5 # Minimum temperature

Air <- TRUE # air heating or floor heating?
Noise <- 0

Stochastic <- FALSE # Stochastic boundaries?
ViolationFraction <- 0.05 # How many temperature violations?
Tmax <- length(data$timedate)-N

#Price
Price <- rep(1,Tmax+N) # Should the prices be constant?
#Price <- prbs(6,322,10)+1 # Or varying?


ForecastedPrice <- Price

source("Setup_model.R")


### Simulate brownian motion

### Simulate measurement errors
Bin <- cumsum(rnorm(Tmax,sd=exp(fit1$xm[8])/fit1$xm[4]*sqrt(dt)))
Bm <- cumsum(rnorm(Tmax,sd=exp(fit1$xm[9])/fit1$xm[5]*sqrt(dt)))

BM <- Noise*rbind(Bin,Bm)

### Simulate the system
Tall <- matrix(0,nrow=2,ncol=Tmax)
u <- numeric(Tmax)


Tall[,1] <- c(23,23)

for(i in 2:Tmax){
  u[i] <- ControlBuilding(c(Price[i],
                            ForecastedPrice[(i+1):(i+N-1)]),
                          c(rbind(data$Ta[i:(i+N-1)],
                                  data$Gv[i:(i+N-1)])),
                          Tall[,(i-1)],Ymax,Ymin,Stochastic)$solution[1]
  Tall[,i] <- Ad%*%Tall[,(i-1)]+1*cBd%*%u[i]+ncBd%*%c(data$Ta[i],data$Gv[i])+t(diff(t(BM)))[,(i-1),drop=FALSE]
}

plot(Tall[1,],type='l',ylim=c(Ymin-0.5,Ymax+0.5),ylab="Air Temperature")
lines(Tall[2,],type='s',col=alpha("blue", 0.4))
lines(u*0.5+Ymin+0.5,type='s',col='red')
lines((Price-1)*0.6+Ymin+1.5,col='green')
lines(c(Ymax,Ymax) ~ c(0,Tmax),lty=2)
lines(c(Ymin,Ymin) ~ c(0,Tmax),lty=2)
legend(70,max(Ymax+0.6),c("Air Temperature","Thermal Mass Temperature",'Heating'),col=c('Black',alpha("blue",0.4),'red'),lty=1,bty='n')

```
\pagebreak