---
title: \Large Excercise 3 - Grey-box models (continued)
author: "Marco Hernandez Velasco"
date: "August 2018"
output: 
  pdf_document:
    citation_package: natbib
header-includes:
  - \usepackage{color}
bibliography: [references.bib, packages.bib]
---
```{r setup_bibliography, include=FALSE}
# automatically create a bib database for R packages
rm(list=ls()) # remove all variables in memory

knitr::write_bib(c(.packages(),  #make a bibliography for the packages
              #packages used that will be cited
                #basic packages used for Rmarkdown
                'knitr', 'rmarkdown',  
                #actual packages used  to solve
                'tidyverse', 'ctsmr', 'digest'),   
                'packages.bib') #file num to which the references will be saved to

#define other general chunk settings
knitr::opts_chunk$set(echo = TRUE, # show source code chunks in the output file
                      fig.height = 4, #figures height for all chunks
                      fig.align = "center") 
```
@* `r #includes all references in file, even those not mentioned in the text `

```{r initialize, message = FALSE, echo = FALSE}
library(readxl) # to import Excel files
library(tidyverse) # to use tidy data
library(ctsmr)
library(digest)

## Init by deleting all variables and functions
rm(list=ls())
## Set the working directory
setwd(".")

## Use the ctsmr package
library(ctsmr)
```

The exercise is focused on grey-box modelling of the heat dynamics of a (small) building using stochastic differential equations (SDEs). In addition to the first exercise on greybox modelling, we will in this exercise test different techniques to:  

1. Alter the noise level or system uncertainty to account for e.g. non-linear phenomena.

2. Build a semi-parametric model to take into account that the solar penetration (i.e. relation between measured solar radiation and radiation entering into the building) as function of the position of the sun.

3. Balance heat gains to the air temperature and the temperature of the thermal mass.  

The data consists of several measurement from a small test box with a single window. In this exercise the following signals are used:

  * Ti (yTi in data) the observed indoor temperatures. (C)  
  * Qi (Qi in data) the heat emitted by the electrical heaters in the test box (W)  
  * Te (Te in data) the ambient temperature (C)  
  * Gv (Gv in the data) the vertical south total solar radiation (W/m2)  
  * Gvn (Gvn in data) the vertical north total solar radiation (W/m2)  

# Question 1

```{r q1_plot, echo=FALSE, warning=FALSE, results='hide'}
## Source functions
sapply(dir("functions",full.names=TRUE), source)

## A list of parameters used in many functions
pg <- list()
## Rerun the models or read from cache?
pg$rerun <- FALSE
pg$cache_dir <- "cache_fits"
## Use first order interpolation when estimating?
pg$firstorder <- TRUE
## Number of threads used, keep at 1
pg$threads <- 1
## Latitude and longitude of the position
pg$latitude <- 37.097083
pg$longitude <- -2.364994

## Averaging period Ts is in minutes
X <- prepare_data(readSeries("X5", Ts=10, nlags=NA), samples_after_Qi_step = 5)

## Plot it
plotmulti(X, c("yTi","Te","Qi","G","stepQi"), xnm="timedate", ylab=c("Ti (c)","Te (C)","Qi (W)","Gv (W/m2)","stepQi"))

```
The lower time series plot is of stepQi, which goes from 0 to 1. Try to change the argument samples_after_Qi_step above in the function preparing the data. How does it change stepQi?

*Making the samples_afterQi_step, makes the on/off intervals longer or shorter and makes it able to correspond better to the actual heater's behavior. Like below with a step = 0.5.*


```{r q1_step_qi_plot, echo=FALSE, warning=FALSE, results='hide'}
## changing samples_after_Qi_step
step_qi <- 0.5

X2 <- prepare_data(readSeries("X5", Ts=10, nlags=NA), samples_after_Qi_step = step_qi)

## Plot it
plotmulti(X2, c("yTi","Te","Qi","G","stepQi"), xnm="timedate", ylab=c("Ti (c)","Te (C)","Qi (W)","Gv (W/m2)","stepQi"))

```
## Comparing the two models

Now compare the two models implemented in functions/sdeTiTw.R and
functions/sdeTiTw_sigmalevels.R. What is the difference?

*The second model includes with sigma levels include*
$(1+(stepQi*sigmalevel))$
*in the error term allowing to account for the noise of turning on/off the heater.*

```{r q1_compare, echo=FALSE, warning=FALSE, results='hide'}
## Keep the fits in a list
Fits <- list()

## TiTw with constant system noise
Fits[["TiTw"]] <- sdeTiTw(X)
## Remember to step back and see the acf and cpgram
analyzeFit(Fits[["TiTw"]], type="s")

## TiTw with two leves of system noise
Fits[["TiTw_sigmalevels"]] <- sdeTiTw_sigmalevels(X)
analyzeFit(Fits[["TiTw_sigmalevels"]], type="s")
```

What is plotted in the upper two plots? (You maybe have to look into the
analyzeFit() function).

*Upper plot is the residuals of the predicted variable yTi*
*lower plot is the standarized residuals (residuals / sd(yTi))*

What is indicated by the blue lines in the upper plot? Step back in the plots and compare the results, and look at the summary output.

*The blue lines in the upper plot are the standard deviation of the predicted variable (yTi) by which the residuals are divided to be standarized*

Which of the two models will you prefer and why?
*The one with sigma level because it allows the variance to change with time and captures better the variability of the process and that is shown in its residuals.*
*When comparing their Log-Likelihood, also the model with sigma-levels has a larger likelihood.*

```{r q1_nice, echo=FALSE, warning=FALSE, results='hide', eval = FALSE}
## Nice features of R and Rstudio  ----------------------------------------

## About keeping the results in a list:
## Easy to extract the same value for all the fits
## E.g. compare the loglikelihoods
sapply(Fits, function(fit){ fit$loglik })

## Shortcuts for running the code in different ways in RStudio:
## See the menu around "Code->Run Region"

## Setting a break point in a function:
## Open "functions/sdeTiTw.R" and click to the left of the "3" number indicating line 3
## A red circle will appear indicating a breakpoint
## Above the script, click on "Source on Save" to check it
## Press the disc sign to save
## Now go back and run the line with sdeTiTw(X)
## The execution stops and you can use the bottoms above the console to step through the code

## Caching results:
## Look into the "functions/TiTw.R" in the first lines the cache_load() is called
## Go into "functions/cache_load.R", can you figure out what it does?
## Now run
sdeTiTw(X, rerun=TRUE)
## and try to change the value of return between TRUE and FALSE
## The results are cached in the folder "cache_fits", such that if a model was previously run on some data, instead of running again, it just opens the saved result from last time

## Keeping results in a list:
## You can easily do the same thing to all the fits, e.g. pick a value or do a plot to sum up results
## E.g. compare the loglikelihoods
sapply(Fits, function(fit){ fit$loglik })
```
\pagebreak